`r params$doc_title`

### 2022-12-03 scripts/MSc.Rmd
mail 2022-11-26

library(dplyr)
library(purrr)
library(tidyr)
library(stringr)
library(tibble)
library(ggplot2)
library(ggthemes)
library(ggpubr)
library(scales)
#library(biomaRt)  # different script
library(kableExtra)
library(MendelianRandomization)
library(TwoSampleMR)
library(mr.raps)
library(MRMix)
library(MRPRESSO)
library(penalized)
library(conflicted)
source("MR_lasso.R")  # Slob&Burgess 2020
source("PVE_calculation_fns.R") 
conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
```

@@

ggplot(data = plot_data_2, aes(x=maf.bmi, y=abs(beta.bmi)))+
  geom_point(alpha=0.6)+
  geom_hline(yintercept=median(abs(plot_data_2$beta.bmi)), 
             col=ggthemes_data$few$colors$Dark[6,2][[1]], alpha=0.6, linetype=2)+
  geom_vline(xintercept=median(plot_data_2$maf.bmi), 
             col=ggthemes_data$few$colors$Dark[6,2][[1]], alpha=0.6, linetype=2)+
  annotate(geom = "text", x = 0.15, y = 0.045, 
           label="relatively rare variants\nwith relatively strong effects", 
           color= ggthemes_data$few$colors$Dark[2,2][[1]]
             )+
  annotate(geom = "text", x = 0.42, y = 0.05, 
           label="relatively common variants\nwith relatively strong effects", 
           color=ggthemes_data$few$colors$Dark[3,2][[1]])+
  annotate(geom = "text", x = 0.1, y = 0.015, 
           label="rare variants\nwith weak effects", color=ggthemes_data$few$colors$Dark[4,2][[1]])+
  annotate("segment", x=0.44,xend=0.42,y=0.025,yend=0.013,arrow=arrow(), 
           color=ggthemes_data$few$colors$Dark[5,2][[1]])+
  annotate(geom = "text", x = 0.44, y = 0.03, 
           label="common variants\nwith weak effects", 
           color=ggthemes_data$few$colors$Dark[5,2][[1]])+
  #geom_text(x=0.1, y=0.05, label="hello", family = "Times New Roman",
  #          size=6, col="red")+
  xlab("Minor allele frequency")+
  ylab("Exposure effect size")+  # abs(beta_BMI)
  labs(title = "Figure 2-update",
       subtitle = "Define different genetic architectures",
       #col='', 
       #caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits explained by the IVs"
       ) +
  theme_few()+
  scale_colour_few() -> fig02.update
fig02.update

@@

  N_annotation <- c(10000, 100000, 200000, 300000)
  ## plot
  plotsubtitle <- paste0("Relatively ", maf ," variants\nwith relatively ", effect_size, " effects")
  ggplot(plot_data1, aes(x=n, y=value, col=name))+
    geom_point(alpha=0.6)+
    geom_line(alpha=0.6)+
    geom_line(aes(x=n, y=count_IVs_scaled, col="Number of IVs"), linetype="dotdash")+
    scale_x_continuous(name = paste0("Size of sample ", as.roman(1)),  # 2-sample-MR or label "Study Size"
                       labels = comma, 
                       breaks = N_annotation, #c( 10000, n[7:length(n)]),
                       limits = c(10000, 300000))+
    #ylab("% variance explained by IVs")+
    labs(title = "Figure 1", subtitle = plotsubtitle,
         col='', 
         caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits   explained by the IVs"
         ) +
    scale_y_continuous(
      # Features of the first axis
      name = "% variance explained by IVs",
      # Add a second axis and specify its features
      sec.axis = sec_axis(~./scale_factor, name="Number of IVs"),
      limits = c(0,2)
    ) + 
    theme_few()+
    #scale_colour_few() +
    #scale_color_manual(name = "", values = c("% var(U)" = "red", 
    #                                         "% var(X)" = "blue", 
    #                                         "% var(Y)" = "darkblue", 
    #                                         "Number of IVs" = "black"
    #                                         ))
    #
    scale_color_manual(name = "", values = c("% var(U)" = ggthemes_data$few$colors$Dark[2,2][[1]], 
                                             "% var(X)" = ggthemes_data$few$colors$Dark[3,2][[1]], 
                                             "% var(Y)" = ggthemes_data$few$colors$Dark[4,2][[1]], 
                                             "Number of IVs" = "black"),
                       # order
                       breaks = c( "Number of IVs", "% var(X)", "% var(Y)", "% var(U)")
    ) -> figure01.quadrant
  print(figure01.quadrant)

@@

```{r fig.height=10, fig.width=14, warning=FALSE}
fig01b.update + 
  labs(title= "Figure 1", caption = "") -> fig01b.update.arrange

fig02.update+
  labs(title= "Genetic architectures", subtitle="", caption = "") -> fig02.update.arrange

figure01_EFFECT_SIZE_strong_VARIANTS_common+
  labs(title= "", caption = "") -> figure01_EFFECT_SIZE_strong_VARIANTS_common.arrange

figure01_EFFECT_SIZE_strong_VARIANTS_rare+
  labs(title= "", caption = "") -> figure01_EFFECT_SIZE_strong_VARIANTS_rare.arrange

figure01_EFFECT_SIZE_weak_VARIANTS_common+
  labs(title= "", caption = "") -> figure01_EFFECT_SIZE_weak_VARIANTS_common.arrange

figure01_EFFECT_SIZE_weak_VARIANTS_rare+
  labs(title= "", caption = "") -> figure01_EFFECT_SIZE_weak_VARIANTS_rare.arrange

ggpubr::ggarrange(fig01b.update.arrange, fig02.update.arrange,
                  figure01_EFFECT_SIZE_strong_VARIANTS_common.arrange, figure01_EFFECT_SIZE_strong_VARIANTS_rare.arrange,
                  figure01_EFFECT_SIZE_weak_VARIANTS_common.arrange, figure01_EFFECT_SIZE_weak_VARIANTS_rare.arrange, 
                  ncol=2, nrow=3, 
                  heights = c(2, 1.5, 1.5),
                  legend = "bottom", common.legend = TRUE, align = "hv"
                  ) -> final_plot1.arrange

@@

ggpubr::ggarrange(ggarrange(fig01b.update.arrange, fig02.update.arrange, ncol=2, labels = c("A", "B"), legend = "none", align = "hv"),
                  ggarrange(figure01_EFFECT_SIZE_strong_VARIANTS_common.arrange, figure01_EFFECT_SIZE_strong_VARIANTS_rare.arrange,
                  figure01_EFFECT_SIZE_weak_VARIANTS_common.arrange, figure01_EFFECT_SIZE_weak_VARIANTS_rare.arrange, 
                  ncol=4, labels = c("C", "D", "E", "F"), legend="bottom", common.legend = TRUE),
                  nrow=2
                  #heights = c(2, 1.5, 1.5),
                  #legend = "bottom", common.legend = TRUE, align = "hv"
                  ) -> final_plot1.arrange2

@@

ggsave(filename = filename_final_plot1, plot = final_plot1.arrange, 
         width = 14, height = 10, 
         units = "in"  # default
         )
ggsave(filename = filename_final_plot1b, plot = final_plot1.arrange2, 
         width = 14, height = 10, 
         units = "in"  # default
         )

>>>>

library(dplyr)
library(purrr)
library(tidyr)
library(stringr)
library(tibble)
library(ggplot2)
library(ggdist)
library(gghalves)
library(ggthemes)
library(ggpubr)
library(scales)
#library(biomaRt)  # different script
library(kableExtra)
library(MendelianRandomization)
library(TwoSampleMR)
library(mr.raps)
library(MRMix)
library(MRPRESSO)
library(penalized)
library(conflicted)
source("MR_lasso.R")  # Slob&Burgess 2020
source("PVE_calculation_fns.R") 
conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
```

@@

ggplot(data = plot_data_2, aes(x=maf.bmi, y=abs(beta.bmi)))+
  geom_point(alpha=0.6)+
  geom_hline(yintercept=median(abs(plot_data_2$beta.bmi)), 
             col=ggthemes_data$few$colors$Dark[6,2][[1]], alpha=0.6, linetype=2)+
  geom_vline(xintercept=median(plot_data_2$maf.bmi), 
             col=ggthemes_data$few$colors$Dark[6,2][[1]], alpha=0.6, linetype=2)+
  annotate(geom = "text", x = 0.15, y = 0.045, 
           label="relatively rare\nrelatively strong", 
           color= ggthemes_data$few$colors$Dark[2,2][[1]]
             )+
  annotate(geom = "text", x = 0.42, y = 0.05, 
           label="relatively common\nrelatively strong", 
           color=ggthemes_data$few$colors$Dark[3,2][[1]])+
  annotate(geom = "text", x = 0.1, y = 0.015, 
           label="relatively rare\nrelatively weak", color=ggthemes_data$few$colors$Dark[4,2][[1]])+
  annotate("segment", x=0.44,xend=0.42,y=0.025,yend=0.013,arrow=arrow(), 
           color=ggthemes_data$few$colors$Dark[5,2][[1]])+
  annotate(geom = "text", x = 0.44, y = 0.03, 
           label="relatively common\nrelatively weak", 
           color=ggthemes_data$few$colors$Dark[5,2][[1]])+
  #geom_text(x=0.1, y=0.05, label="hello", family = "Times New Roman",
  #          size=6, col="red")+
  xlab("Minor allele frequency")+
  ylab("Exposure effect size")+  # abs(beta_BMI)
  labs(title = "Figure 2-update",
       subtitle = "Define different genetic architectures",
       #col='', 
       #caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits explained by the IVs"
       ) +
  theme_few()+
  scale_colour_few() -> fig02.update
fig02.update

@@

  N_annotation <- c(10000, 100000, 200000, 300000)
  ## plot
  plotsubtitle <- paste0("relatively ", maf ,", relatively ", effect_size)
  ggplot(plot_data1, aes(x=n, y=value, col=name))+
    geom_point(alpha=0.6)+
    geom_line(alpha=0.6)+
    geom_line(aes(x=n, y=count_IVs_scaled, col="Number of IVs"), linetype="dotdash")+
    scale_x_continuous(name = paste0("Size of sample ", as.roman(1)),  # 2-sample-MR or label "Study Size"
                       labels = comma, 
                       breaks = N_annotation, #c( 10000, n[7:length(n)]),
                       limits = c(10000, 300000))+
    #ylab("% variance explained by IVs")+
    labs(title = "Figure 1", subtitle = plotsubtitle,
         col='', 
         caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits   explained by the IVs"
         ) +
    scale_y_continuous(
      # Features of the first axis
      name = "% variance explained by IVs",
      # Add a second axis and specify its features
      sec.axis = sec_axis(~./scale_factor, name="Number of IVs"),
      limits = c(0,2)
    ) + 
    theme_few()+
    #scale_colour_few() +
    #scale_color_manual(name = "", values = c("% var(U)" = "red", 
    #                                         "% var(X)" = "blue", 
    #                                         "% var(Y)" = "darkblue", 
    #                                         "Number of IVs" = "black"
    #                                         ))
    #
    scale_color_manual(name = "", values = c("% var(U)" = ggthemes_data$few$colors$Dark[2,2][[1]], 
                                             "% var(X)" = ggthemes_data$few$colors$Dark[3,2][[1]], 
                                             "% var(Y)" = ggthemes_data$few$colors$Dark[4,2][[1]], 
                                             "Number of IVs" = "black"),
                       # order
                       breaks = c( "Number of IVs", "% var(X)", "% var(Y)", "% var(U)")
    ) -> figure01.quadrant
  print(figure01.quadrant)

@@

```{r fig.height=10, fig.width=14, warning=FALSE}
fig01b.update + 
  labs(title= "", caption = "") -> fig01b.update.arrange

fig02.update+
  labs(title= "", subtitle="", caption = "") -> fig02.update.arrange

figure01_EFFECT_SIZE_strong_VARIANTS_common+
  labs(title= "", caption = "") +
  theme(plot.subtitle = element_text(color = ggthemes_data$few$colors$Dark[3,2][[1]])) -> figure01_EFFECT_SIZE_strong_VARIANTS_common.arrange

figure01_EFFECT_SIZE_strong_VARIANTS_rare+
  labs(title= "", caption = "") +
  theme(plot.subtitle = element_text(color = ggthemes_data$few$colors$Dark[2,2][[1]])) -> 
  figure01_EFFECT_SIZE_strong_VARIANTS_rare.arrange

figure01_EFFECT_SIZE_weak_VARIANTS_common+
  labs(title= "", caption = "") +
  theme(plot.subtitle = element_text(color = ggthemes_data$few$colors$Dark[5,2][[1]])) -> figure01_EFFECT_SIZE_weak_VARIANTS_common.arrange

figure01_EFFECT_SIZE_weak_VARIANTS_rare+
  labs(title= "", caption = "") +
  theme(plot.subtitle = element_text(color = ggthemes_data$few$colors$Dark[4,2][[1]])) -> figure01_EFFECT_SIZE_weak_VARIANTS_rare.arrange

ggpubr::ggarrange(fig02.update.arrange, fig01b.update.arrange,
                  figure01_EFFECT_SIZE_strong_VARIANTS_common.arrange, figure01_EFFECT_SIZE_strong_VARIANTS_rare.arrange,
                  figure01_EFFECT_SIZE_weak_VARIANTS_common.arrange, figure01_EFFECT_SIZE_weak_VARIANTS_rare.arrange, 
                  ncol=2, nrow=3, 
                  heights = c(2, 1.5, 1.5),
                  legend = "bottom", common.legend = TRUE, align = "hv"
                  ) -> final_plot1.arrange

@@

ggpubr::ggarrange(ggarrange(fig02.update.arrange, fig01b.update.arrange, ncol=2, labels = c("A", "B"), legend = "none", align = "hv"),
                  ggarrange(figure01_EFFECT_SIZE_strong_VARIANTS_rare.arrange,
                            figure01_EFFECT_SIZE_strong_VARIANTS_common.arrange, 
                            figure01_EFFECT_SIZE_weak_VARIANTS_rare.arrange,
                            figure01_EFFECT_SIZE_weak_VARIANTS_common.arrange, 
                            ncol=4, labels = c("C", "D", "E", "F"), 
                            legend="bottom", common.legend = TRUE),
                  nrow=2
                  #heights = c(2, 1.5, 1.5),
                  #legend = "bottom", common.legend = TRUE, align = "hv"
                  ) -> final_plot1.arrange2

@@

ggsave(filename = filename_final_plot1, plot = final_plot1.arrange, 
         width = 14, height = 10, 
         units = "in"  # default
         )
ggsave(filename = filename_final_plot1b, plot = final_plot1.arrange2, 
         width = 14, height = 10, 
         units = "in"  # default
         )
```
